Temperature and wave exposure as drivers of morphology in two kelp species

### 2019-07-15 paper_morph_new.Rmd
morphology graphs for laminaria

---
title: Temperature, wave exposure and wind as drivers of morphology in two kelp species
  along the coast of South Africa
author: "R. Coppin, C. Rautenbach, T. Ponton and A.J Smit"
geometry: margin=1in
output:
  pdf_document: default
  word_document:
    fig_caption: yes
papersize: A4
tables: yes
fontsize: 10pt
---

@@

# Results
## Temperature variables

```{r Temperature data prep, message=FALSE, warning=FALSE, include=FALSE}

library(tidyverse)
library(ggpubr)
library(gridExtra)

## Preparing data
# Complete lon/lats
kelp_sites <- read.csv("metadata/site_meta_complete.csv", sep = ";")

@@

## Morphology
### *L. pallida*
```{r laminaria processing, message=FALSE, warning=FALSE, include=FALSE}

lam_data <- read.csv("data/morph_data/lam_morph_data.csv", sep = ";") %>% 
   filter(site!= "Rocky Bank")

@@

pos_lam <- c("Betty's Bay", "Roman Rock","A-Frame", "Miller's Point", "Baboon Rock", 
                  "Batsata Rock", "Buffels Bay"," Buffels Bay South","Bordjies North","Olifantsbos", "Kommetjie", "Sea Point")
pos_exposure_lam <- c("Fully sheltered", "Extremely sheltered", "Very sheltered", "Sheltered", "Moderatley exposed", "Exposed", "Very exposed")

lamina_weight_plot <- ggplot(data = lam_df, aes(x = wave_exposure, y = laminaWeight, fill = wave_exposure)) +
         scale_x_discrete(limits = pos_exposure_lam) + 
         geom_boxplot(outlier.shape = NA) 
lamina_length_plot <- ggplot(data = lam_df, aes(x = wave_exposure, y = laminaLength, fill = wave_exposure)) +
         scale_x_discrete(limits = pos_exposure_lam) + 
         geom_boxplot(outlier.shape = NA) 
lamina_thickness_plot <- ggplot(data = lam_df, aes(x = wave_exposure, y = laminaThickness, fill = wave_exposure)) +
         scale_x_discrete(limits = pos_exposure_lam) + 
         geom_boxplot(outlier.shape = NA) 
stipe_mass_lam_plot <- ggplot(data = lam_df, aes(x = wave_exposure, y = stipeMass, fill = wave_exposure)) +
         scale_x_discrete(limits = pos_exposure_lam) + 
         geom_boxplot(outlier.shape = NA) 
stipe_length_lam_plot <- ggplot(data = lam_df, aes(x = wave_exposure, y = stipeLength, fill = wave_exposure)) +
         scale_x_discrete(limits = pos_exposure_lam) + 
         geom_boxplot(outlier.shape = NA) 
stipe_diameter_plot <- ggplot(data = lam_df, aes(x = wave_exposure, y = stipeDiameter, fill = wave_exposure)) +
         scale_x_discrete(limits = pos_exposure_lam) + 
         geom_boxplot(outlier.shape = NA) 
digits_plot <- ggplot(data = lam_df, aes(x = wave_exposure, y = digits, fill = wave_exposure)) +
         scale_x_discrete(limits = pos_exposure_lam) + 
         geom_boxplot(outlier.shape = NA) 
thallus_mass_plot <- ggplot(data = lam_df, aes(x = wave_exposure, y = thallusMass, fill = wave_exposure)) +
         scale_x_discrete(limits = pos_exposure_lam) + 
         geom_boxplot(outlier.shape = NA) 
total_length_plot <- ggplot(data = lam_df, aes(x = wave_exposure, y = totalLength, fill = wave_exposure)) +
         scale_x_discrete(limits = pos_exposure_lam) + 
         geom_boxplot(outlier.shape = NA) 
```

```{r laminaria plots, echo=FALSE, message=FALSE, warning=FALSE}
grid.arrange(lamina_weight_plot,lamina_length_plot,lamina_thickness_plot,stipe_mass_lam_plot,stipe_length_lam_plot,stipe_diameter_plot, digits_plot, thallus_mass_plot,total_length_plot,total_length_plot, ncol = 3, nrow = 4)

@@

>>>>

---
title: Temperature and wave exposure as drivers of morphology in two kelp species
  along the coast of South Africa
author: "R. Coppin, C. Rautenbach, T. Ponton and A.J Smit"
geometry: margin=1in
output:
  word_document:
    fig_caption: yes
  pdf_document: default
papersize: A4
tables: yes
fontsize: 10pt
---

@@

# Results
## Temperature variables
```{r software packages, message=TRUE, warning=TRUE, include=FALSE, echo=FALSE}
library(tidyverse)
library(ggpubr)
library(gridExtra)
```

```{r Temperature data prep, message=FALSE, warning=FALSE, include=FALSE}
## Data wrangling and preparation
## Preparing data
# Complete lon/lats
kelp_sites <- read.csv("metadata/site_meta_complete.csv", sep = ";")

@@

## Morphology
### *L. pallida*

```{r laminaria processing, message=FALSE, warning=FALSE, include=FALSE}
lam_data <- read.csv("data/morph_data/lam_morph_data.csv", sep = ";") %>% 
   filter(site!= "Rocky Bank")

@@

pos_lam_temp <- c("Cool temperate", "Warm temperate")

LT1 <- ggplot(data = lam_df, aes(x = temp_regime, y = laminaWeight, fill = temp_regime)) +
  scale_x_discrete(limits = pos_lam_temp) + 
  geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE) + 
  stat_compare_means() + 
  xlab("") + 
  ylab("Lamina weight") +
  scale_fill_discrete(name="Temperature regime") +
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank())
LT2 <- ggplot(data = lam_df, aes(x = temp_regime, y = laminaLength, fill = temp_regime)) +
  scale_x_discrete(limits = pos_lam_temp)  + 
  geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE) + 
  stat_compare_means() + 
  xlab("") + 
  ylab("Lamina length") +
  scale_fill_discrete(name="Temperature regime") +
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank())
LT3 <- ggplot(data = lam_df, aes(x = temp_regime, y = laminaThickness, fill = temp_regime)) +
  scale_x_discrete(limits = pos_lam_temp) + 
  geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE) + 
  stat_compare_means() + 
  xlab("") + 
  ylab("Lamina thickness") +
  scale_fill_discrete(name="Temperature regime") +
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank())
LT4 <- ggplot(data = lam_df, aes(x = temp_regime, y = stipeMass, fill = temp_regime)) +
  scale_x_discrete(limits = pos_lam_temp)  + 
  geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE) + 
  stat_compare_means() + 
  xlab("") + 
  ylab("Stipe mass (g)") +
  scale_fill_discrete(name="Temperature regime") +
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank())
LT5 <- ggplot(data = lam_df, aes(x = temp_regime, y = stipeLength, fill = temp_regime)) +
  scale_x_discrete(limits = pos_lam_temp)  + 
  geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE) + 
  stat_compare_means() + 
  xlab("") + 
  ylab("Stipe length (cm)") +
  scale_fill_discrete(name="Temperature regime") +
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank())
LT6 <- ggplot(data = lam_df, aes(x = temp_regime, y = stipeDiameter, fill = temp_regime)) +
  scale_x_discrete(limits = pos_lam_temp)  + 
  geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE) + 
  stat_compare_means() + 
  xlab("") + 
  ylab("Stipe diameter (cm)") +
  scale_fill_discrete(name="Temperature regime") +
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank())
LT7 <- ggplot(data = lam_df, aes(x = temp_regime, y = digits, fill = temp_regime)) +
  scale_x_discrete(limits = pos_lam_temp)  + 
  geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE) + 
  stat_compare_means() + 
  xlab("") + 
  ylab("No. of digits") +
  scale_fill_discrete(name="Temperature regime") +
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank())
LT8 <- ggplot(data = lam_df, aes(x = temp_regime, y = thallusMass, fill = temp_regime)) +
  scale_x_discrete(limits = pos_lam_temp) + 
  geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE) + 
  stat_compare_means() + 
  xlab("") + 
  ylab("Thallus mass (g)") +
  scale_fill_discrete(name="Temperature regime") +
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank())
LT9 <- ggplot(data = lam_df, aes(x = temp_regime, y = totalLength, fill = temp_regime)) +
  scale_x_discrete(limits = pos_lam_temp) + 
  geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE) + 
  stat_compare_means() + 
  xlab("") + 
  ylab("Total length (cm)") +
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank())

lam_df$wave_exposure <- factor(lam_df$wave_exposure, levels=c("Very sheltered", "Very exposed", "Moderately exposed","Fully sheltered","Extremely sheltered"), labels=c("Fully sheltered", "Extremely sheltered", "Very sheltered", "Moderately exposed", "Very exposed"))
pos_exposure_lam <- c("Fully sheltered", "Extremely sheltered", "Very sheltered", "Moderately exposed", "Very exposed")

LW1 <- ggplot(data = lam_df, aes(x = wave_exposure, y = laminaWeight, fill = wave_exposure)) +
  scale_x_discrete(limits = pos_exposure_lam) + 
  geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = TRUE) +
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank()) +
  ylab("Lamina weight (g)") + 
  scale_fill_brewer(palette="OrRd") +
  guides(fill=guide_legend(title="Wave exposure level"))

LW2 <- ggplot(data = lam_df, aes(x = wave_exposure, y = laminaLength, fill = wave_exposure)) +
  scale_x_discrete(limits = pos_exposure_lam)  + 
  geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = TRUE) +
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank()) +
  ylab("Lamina length (cm)") + 
  scale_fill_brewer(palette="OrRd") +
  guides(fill=guide_legend(title="Wave exposure level"))
LW3 <- ggplot(data = lam_df, aes(x = wave_exposure, y = laminaThickness, fill = wave_exposure)) +
  scale_x_discrete(limits = pos_exposure_lam) + 
  geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = TRUE) +
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank()) +
  ylab("Lamina thickness") + 
  scale_fill_brewer(palette="OrRd") +
  guides(fill=guide_legend(title="Wave exposure level")) 
LW4 <- ggplot(data = lam_df, aes(x = wave_exposure, y = stipeMass, fill = wave_exposure)) +
  scale_x_discrete(limits = pos_exposure_lam)  + 
  geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = TRUE) +
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank()) +
  ylab("Stipe mass (g)") + 
  scale_fill_brewer(palette="OrRd") +
  guides(fill=guide_legend(title="Wave exposure level")) 
LW5 <- ggplot(data = lam_df, aes(x = wave_exposure, y = stipeLength, fill = wave_exposure)) +
  scale_x_discrete(limits = pos_exposure_lam)  + 
  geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = TRUE) +
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank()) +
  ylab("Stipe length (cm)") + 
  scale_fill_brewer(palette="OrRd") +
  guides(fill=guide_legend(title="Wave exposure level")) 
LW6 <- ggplot(data = lam_df, aes(x = wave_exposure, y = stipeDiameter, fill = wave_exposure)) +
  scale_x_discrete(limits = pos_exposure_lam)  + 
  geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = TRUE) +
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank()) +
  ylab("Stipe diameter (cm)") + 
  scale_fill_brewer(palette="OrRd") +
  guides(fill=guide_legend(title="Wave exposure level")) 
LW7 <- ggplot(data = lam_df, aes(x = wave_exposure, y = digits, fill = wave_exposure)) +
  scale_x_discrete(limits = pos_exposure_lam)  + 
  geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = TRUE) +
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank()) +
  ylab("No. of digits") + 
  scale_fill_brewer(palette="OrRd") +
  guides(fill=guide_legend(title="Wave exposure level")) 
LW8 <- ggplot(data = lam_df, aes(x = wave_exposure, y = thallusMass, fill = wave_exposure)) +
  scale_x_discrete(limits = pos_exposure_lam) + 
  geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = TRUE) +
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank()) +
  ylab("Thallus mass (kg)") + 
  scale_fill_brewer(palette="OrRd") +
  guides(fill=guide_legend(title="Wave exposure level")) 
LW9 <- ggplot(data = lam_df, aes(x = wave_exposure, y = totalLength, fill = wave_exposure)) +
  scale_x_discrete(limits = pos_exposure_lam) + 
  geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = TRUE) +
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank()) +
  ylab("Total length (cm)") + 
  scale_fill_brewer(palette="OrRd") +
  guides(fill=guide_legend(title="Wave exposure level")) 

@@

```{r laminaria temp plots, echo=FALSE, message=FALSE, warning=FALSE}
ggarrange(LT1, LT2,LT3,LT4,LT5,LT6,LT7,LT8,LT9, labels = c('A', 'B', 'C','D','E','F','G','H',"I"), common.legend = TRUE, legend = "bottom", label = "p.format", ncol = 3, nrow = 3)
```

```{r laminaria wave plots, echo=FALSE, message=FALSE, warning=FALSE}
ggarrange(LW1, LW2,LW3,LW4,LW5,LW6,LW7,LW8,LW9, labels = c('A', 'B', 'C','D','E','F','G','H',"I"), common.legend = TRUE, legend = "bottom")
```
